---
title: "Mapquery_scRNA_eR"
author: "Huixin Jin"
date: "2023-03-21"
output: html_document
---

```{r}
library(Seurat)
library(Signac)
library(EnsDb.Hsapiens.v86)
library(ggplot2)
library(cowplot)
library(reticulate)
library(SingleCellExperiment)
library(tidyverse)
library(RColorBrewer)
library(scCustomize)
library(ggrepel)
library(aplot)

set.seed(123)
```

# scRNA process
## preprocess (filter, normalization, regress out)
```{r, message=FALSE}
# read in scRNA data and process
## current work directory "/Users/jinhuixin/Master/thesis/BS1140"
bs1140.rna <- Read10X('/Users/jinhuixin/Master/thesis/aligned_raw_data/BS1140/outs/filtered_feature_bc_matrix/')
bs1140.rna <- bs1140.rna$`Gene Expression`
geneTCRBCR <- function(Data = covidold1798.data){
  TCRBCR.features <- c(grep('^TRAV', rownames(Data)),
                       grep('^TRAJ', rownames(Data)),
                       grep('^TRBV', rownames(Data)),
                       grep('^TRBD', rownames(Data)),
                       grep('^TRBJ', rownames(Data)))
  return(TCRBCR.features)
}
features_to_delete <- geneTCRBCR(Data = bs1140.rna)
bs1140.rna <- bs1140.rna[-features_to_delete,] 
```

```{r}
bs1140.seurat <- CreateSeuratObject(counts = bs1140.rna, project = "bs1140", min.cells = 3, min.features = 200) # 5301-->5292
table(bs1140.seurat$orig.ident)

bs1140.seurat[["percent.mt"]] <- PercentageFeatureSet(object = bs1140.seurat, pattern = "^MT-")
bs1140.seurat[["percent.hsp"]] <- PercentageFeatureSet(object = bs1140.seurat, pattern = "^HSP")

VlnPlot(bs1140.seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(object = bs1140.seurat, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object = bs1140.seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
```

```{r}
#nFeature_RNA: genes per cell; nCount_RNA: molecules detected per cell
bs1140.seurat <- subset(x = bs1140.seurat, subset = nFeature_RNA > 150 & nFeature_RNA < 3000 & percent.mt < 10) 

#Normalize each dataset via scTransform (fast and accurate, may want to run it on the cluster)
#Regress out MT and HSP genes
bs1140.seurat <- SCTransform(bs1140.seurat, vars.to.regress = c("percent.mt", "percent.hsp"), verbose = FALSE)

bs1140.seurat <- RunPCA(bs1140.seurat, features = VariableFeatures(bs1140.seurat), assay = "SCT", npcs = 50)
DimPlot(bs1140.seurat, reduction = 'pca')
bs1140.seurat <- RunUMAP(object = bs1140.seurat, assay = "SCT", reduction = "pca", return.model = TRUE, dims = 1:50)
bs1140.seurat <- FindNeighbors(object = bs1140.seurat, assay = "SCT", reduction = "pca", dims = 1:50)
bs1140.seurat <- FindClusters(object = bs1140.seurat, resolution = 1)
```

```{r}
DefaultAssay(bs1140.seurat) <- 'SCT'
# find markers
markers <- FindAllMarkers(bs1140.seurat, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
markers %>%
  group_by(cluster) %>%
  slice_max(n = 10, order_by = avg_log2FC)
```

SCT |Markers |Cell Type
-----|-----|----- 
0|ENTPD1|Ex-ENTPD1+
1|GZMK|M-GZMK+
2|IL7R|M-IL7R+
3|CD69|M-CD69+
4|GNLY, NKG7|GNLY+NKG7+
5|SESN3|Naive-SESN3+
6|CMSS1, FILIP1L|CMSS1+FILIP1L+
7|IKZF2, PITPNC1|IKZF2+PITPNC1+
8|HIST1H4C, STMN1|Proliferate-STMN1+TOP2A
9|GNLY, IKZF2|GNLY+IKZF2+

```{r}
# assign cell type on each cluster
new.cluster.ids <- c('Ex-ENTPD1+', 'M-GZMK+', 'M-IL7R+', 'M-CD69+', 'GNLY+NKG7+', 'Naive-SESN3+', 'CMSS1+FILIP1L+', 'IKZF2+PITPNC1+', 'Prol-STMN1+TOP2A+', 'GNLY+IKZF2+')
names(new.cluster.ids) <- levels(bs1140.seurat)
bs1140.seurat <- RenameIdents(bs1140.seurat, new.cluster.ids)
bs1140.seurat<- AddMetaData(bs1140.seurat, Idents(bs1140.seurat), col.name = "celltype_rna_sct_assay")
DimPlot(bs1140.seurat, reduction = 'umap', label = TRUE) + NoLegend()
```

# eR adata
```{r}
# set environment
use_condaenv('env_x86')
```

```{r}
cell_order <- colnames(bs1140.seurat)
cell_order <- paste0(cell_order, '-BS1140')
```


```{python}
import scanpy as sc
# eR
adata = sc.read_h5ad('/Users/jinhuixin/Master/thesis/GRN/BS1140/eRegulon_cluster/eR_cluster_anndata.h5ad')
adata = adata[r.cell_order,]
```


## Create a Seurat object
```{r}
# Get the expression matrix
exprs <- t(py$adata$X)
colnames(exprs) <- py$adata$obs_names$to_list()
rownames(exprs) <- py$adata$var_names$to_list()
# Create the Seurat object
eR.seurat <- CreateSeuratObject(exprs)
# Set the expression assay
eR.seurat <- SetAssayData(eR.seurat, "data", exprs)
# Add observation metadata
eR.seurat <- AddMetaData(eR.seurat, py$adata$obs)
# Add feature metadata
#eR.seurat[["RNA"]][["n_cells"]] <- py$adata$var["n_cells"] 
# Add embedding
embedding <- py$adata$obsm["X_umap"] # umap based on eR
rownames(embedding) <- py$adata$obs_names$to_list()
colnames(embedding) <- c("umap_1","umap_2")
eR.seurat[["umap"]] <- CreateDimReducObject(embedding, key = "umap_")

embedding_pca <- py$adata$obsm["X_pca"] # pca based on eR
rownames(embedding_pca) <- py$adata$obs_names$to_list()
colnames(embedding_pca) <- paste0(rep("PC_",50),c(1:50))
eR.seurat[["pca"]] <- CreateDimReducObject(embedding_pca, key = "PC_")

eR.seurat <- SetIdent(eR.seurat, value = 'leiden')
```

## show RNA and eRegulon umap together
```{r}
p1 <- DimPlot(bs1140.seurat) + NoLegend() + ggtitle('RNA')
p2 <- DimPlot(eR.seurat, reduction = 'umap') + NoLegend() + ggtitle('eRegulon')
p1 + p2
p2
```

# build a new combined seurat object
```{r}
n.bs1140 <- bs1140.seurat
newnames <- gsub('-BS1140', '', colnames(eR.seurat))
eR.seurat <- RenameCells(eR.seurat, new.names = newnames)
eRassay <- GetAssayData(eR.seurat, assay = 'RNA', slot = "data")
# subsample common cells
n.bs1140<- subset(n.bs1140, cells = colnames(eR.seurat))
n.bs1140[['eR']] <- CreateAssayObject(data = eRassay)
```

```{r}
# dimention reduction on eR
DefaultAssay(n.bs1140) <- "eR"
VariableFeatures(n.bs1140) <- rownames(n.bs1140[['eR']]) # use all eR features for dimensional reduction
n.bs1140@reductions[['pca_eR']] <- eR.seurat@reductions[['pca']]
n.bs1140@reductions[['umap_eR']] <- eR.seurat@reductions[['umap']]
n.bs1140@reductions[['pca_rna']] <- n.bs1140@reductions[['pca']]
n.bs1140@reductions[['pca']] <- NULL
n.bs1140@reductions[['umap_rna']] <- n.bs1140@reductions[['umap']]
n.bs1140@reductions[['umap']] <- NULL
n.bs1140 <-  ScaleData(n.bs1140) %>%RunPCA(reduction.name = 'pca')

n.bs1140 <- RunUMAP(n.bs1140, reduction.name = 'umap', dims = 1:50, reudction = 'pca', return.model = TRUE)
DimPlot(n.bs1140, reduction = 'umap', label = TRUE)
```

## add the eR cluster cell type info
```{r}
tmp <- read.csv('/Users/jinhuixin/Master/thesis/GRN/BS1140/cluster_overlap/eRegulon_celltype.csv', row.names = 1)
rownames(tmp) <- gsub('-BS1140','',rownames(tmp))
reorder_tmp <- tmp[rownames(n.bs1140@meta.data),]
n.bs1140@meta.data[['celltype_eR']] <- reorder_tmp
rm(tmp, reorder_tmp)
```

## add ATAC cell type info to see the difference in 0 and 2 in wnn
```{r}
n.bs1140@meta.data[['celltype_atac']] <- eR.seurat@meta.data[rownames(n.bs1140@meta.data),'ACC_celltype']
```


# FindTransferAnchor on BS1140 5'RNA-only data
label bs1140 5'scRNA dataset via mapquery to the integrated bs1140 multiomic object

```{r}
bs1140.5rna <- readRDS('/Users/jinhuixin/Master/thesis/5primeRNA_data/bs1140_5primeRNAonly.rds') 
```

## compute transfer anchor
```{r}
transfer.anchors2 <- FindTransferAnchors(reference = n.bs1140, 
                                        query = bs1140.5rna, 
                                        features = VariableFeatures(n.bs1140, assay = 'SCT'),
                                        reference.assay = 'SCT',
                                        query.assay = 'SCT',
                                        normalization.method = 'SCT',
                                        reduction = 'cca')

# predict rna cell type
celltype.predictions2_rna <- TransferData(anchorset = transfer.anchors2,
                                     refdata = n.bs1140$celltype_rna_sct_assay,
                                     weight.reduction = bs1140.5rna[['umap']], 
                                     dims = 1:2)
colnames(celltype.predictions2_rna)[1] <- 'predicted.id.rna' 
bs1140.5rna <- AddMetaData(bs1140.5rna, metadata = celltype.predictions2_rna, col.name = colnames(celltype.predictions2_rna))

# predict eR cell type
celltype.predictions2_eR <- TransferData(anchorset = transfer.anchors2,
                                     refdata = n.bs1140$celltype_eR,
                                     weight.reduction = bs1140.5rna[['umap']], 
                                     dims = 1:2)
colnames(celltype.predictions2_eR)[1] <- 'predicted.id.eR'
bs1140.5rna <- AddMetaData(bs1140.5rna, metadata = celltype.predictions2_eR)

# predict atac cell type 
celltype.predictions2_atac <- TransferData(anchorset = transfer.anchors2,
                                     refdata = n.bs1140$celltype_atac,
                                     weight.reduction = bs1140.5rna[['umap']], 
                                     dims = 1:2)
colnames(celltype.predictions2_atac)[1] <- 'predicted.id.atac'
bs1140.5rna <- AddMetaData(bs1140.5rna, metadata = celltype.predictions2_atac)
```

## UMAP projection to eR umap
### compute the model from scanpy eR cluster
```{r}
# try use seurat umap to get a model
eR.forumap <- eR.seurat
DefaultAssay(eR.forumap) <- 'RNA'
VariableFeatures(eR.forumap) <- rownames(eR.forumap)
eR.forumap <- RunUMAP(eR.forumap,
                      return.model = TRUE, 
                      umap.method = 'umap-learn',
                      n.neighbors = 15,
                      min.dist = 0.1,
                      n.components = 2,
                      metric = 'euclidean',
                      features = VariableFeatures(eR.forumap),
                      seed.use = 555) 
```

```{r}
# replace old eR umap
n.bs1140@reductions[['umap_eR']] <- eR.forumap@reductions[['umap']]
n.bs1140@reductions[['pca_eR']] <- eR.forumap@reductions[['pca']]

bs1140.5rna <- MapQuery(anchorset = transfer.anchors2, 
                        reference = n.bs1140, 
                        query = bs1140.5rna,
                        reference.reduction = 'umap_eR',
                        reduction.model = 'umap_eR')

my_col_pal2 <- c("M-GZMK+"="#8DD3C7", "Ex-ENTPD1+"="#FFFFB3", "M-IL7R+"="#BEBADA", "GNLY+NKG7+"="#FB8072", "M-CD69+"="#80B1D3", "Naive-SESN3+"="#FDB462", "IKZF2+PITPNC1+"="#B3DE69", "Prol-STMN1+TOP2A+"="#FCCDE5", "GNLY+IKZF2+"="#D9D9D9", "CMSS1+FILIP1L+" = "#BC80BD")

p1 <- DimPlot(n.bs1140, reduction = 'umap_eR', group.by = 'celltype_rna_sct_assay', label = TRUE, label.size = 3, repel = TRUE, cols = my_col_pal2) + 
  ggtitle('Reference RNA SCT annotations')
p2 <- DimPlot(bs1140.5rna, reduction = 'ref.umap', group.by = 'predicted.id.rna', label = TRUE, label.size = 3, repel = TRUE, cols = my_col_pal2, pt.size = 0.7) + 
  NoLegend() + 
  ggtitle('Query transferred RNA SCT labels')
p3 <- DimPlot(n.bs1140, reduction = 'umap_eR', group.by = 'celltype_eR', label = TRUE, label.size = 3, repel = TRUE, cols = my_color) + 
  ggtitle('Reference eR annotations')
p4 <- DimPlot(bs1140.5rna, reduction = 'ref.umap', group.by = 'predicted.id.eR', label = TRUE, label.size = 3, repel = TRUE, cols = my_color, pt.size = 0.7) + 
  NoLegend() + 
  ggtitle('Query transferred eR labels')
p5 <- DimPlot(n.bs1140, reduction = 'umap_eR', group.by = 'celltype_atac', label = TRUE, label.size = 3, repel = TRUE, cols = my_color) + 
  ggtitle('Reference eR annotations')
p6 <- DimPlot(bs1140.5rna, reduction = 'ref.umap', group.by = 'predicted.id.atac', label = TRUE, label.size = 3, repel = TRUE, cols = my_color, pt.size = 0.7) + 
  NoLegend() + 
  ggtitle('Query transferred eR labels')
p <- (p1 / p2) | (p3 / p4) | (p5 / p6)
pdf('./MapQuery_bs1140_5p_rna_on_eR_umap.pdf', width = 24, height = 8)
p
dev.off()
```

# save mapquery results of 5'rna
```{r}
saveRDS(bs1140.5rna, file = './bs1140_5prime_rna_mapquery_seurat_obj.rds')
```

# save bs1140 reference object
```{r}
saveRDS(n.bs1140, file = './bs1140_rna_eR_integrated_seurat_obj.rds')
```


